---
title: "Project 3 - Early"
author: "Joel Revo, Kenny Andrysiak, Jacob Cohen, Alec Palo, Michael McLaughlan"
date: "3/29/2021"
output: 
  html_document: 
    code_folding: hide
    number_sections: true 
    toc: true
    toc_float: 
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
library(VGAM)
library(class)
library(caret)
library(gmodels)
library(tidyverse)
library(lmtest)
library(aod)
library(neuralnet)
library(C50)
```
#Abstract and Guiding Question 

Dataset 
(Assumption) We are assuming that the dataset is taken from many different passengers from different airlines. However, we do not have data on each individual airline. So our audience will be the Airline industry as a whole. We will tailor our results to different airline categories (Luxury, standard, budget). 

Guiding Question 

1. We want to predict whether or not a given customer will be satisfied or not

2. What parts of the customer experience contribute most to whether a customer is satisfied or not, and how would differnt airlines best implement their services to make a customer satisfied? 


#Data Cleaning 
```{r}

Airlinedata<-read.csv("Invistico_Airline.csv")


Airlinedata$Gender<-as.factor(Airlinedata$Gender)
Airlinedata$Customer.Type<-as.factor(Airlinedata$Gender)
Airlinedata$Type.of.Travel<-as.factor(Airlinedata$Type.of.Travel)
Airlinedata$Class<-as.factor(Airlinedata$Class)



Airlinedata$satisfaction<-ifelse(Airlinedata$satisfaction=="satisfied", 1, 0)

Airlinedata <- na.omit(Airlinedata)

summary(Airlinedata)
str(Airlinedata)



```

#Normalize The Data
```{r}
Airlinemm <- as.data.frame(model.matrix(~.-1,Airlinedata))


# Randomize the rows in the data (shuffling the rows)
set.seed(12345)
airline_random <- Airlinemm[sample(nrow(Airlinemm)),]

#Normalize the data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# we are going to normalize everything 
airline_norm <- as.data.frame(lapply(airline_random, normalize))
```
#Get Training and Testing Samples
```{r}
airline_train <- airline_norm[1:105000,]
airline_test <- airline_norm[105001:129487,]


airline_train_labels <- airline_norm[1:105000, "satisfaction"]
airline_test_labels <- airline_norm[105001:129487, "satisfaction"]
```

#Logistic Model 
```{r cache=TRUE}

airline_glm <- glm(satisfaction ~ ., data = airline_train, family= "binomial")

summary(airline_glm)

step_airline<-step(airline_glm)
```

#Step Model Summary
```{r}
summary(step_airline)
```
##Logistic Model Predictions 
```{r}
AirlineGLM_pred <- ifelse(predict(step_airline, newdata =airline_test,  type = "response") >= .5, 1, 0)

CrossTable(airline_test$satisfaction, AirlineGLM_pred)

#GETTING ERROR THAT there is an unused argument here. However, you aboslutely need positive =1, otherwise the results flip. Not sure what that error is. 

#confusionMatrix(as.factor(AirlineGLM_pred), as.factor(airline_test$satisfaction), positive = "1")
```

## Logistic Model (Optimal Threshold)
```{r}
library(InformationValue)

#Creating Predictions (percentage)
AirlineGLM_pred_optimal <- predict(step_airline, newdata =airline_test,  type = "response")

#Finding optimal threshold
optimalGLM<-optimalCutoff(airline_test$satisfaction, AirlineGLM_pred_optimal)[1]

#if predictions> optimal, 1, otherwise 0
Airline_binary_predictions<-ifelse(AirlineGLM_pred_optimal>optimalGLM,1,0)


CrossTable(airline_test$satisfaction, Airline_binary_predictions)

#GETTING ERROR THAT there is an unused argument here. However, you aboslutely need positive =1, otherwise the results flip. Not sure what that error is. 

#confusionMatrix(as.factor(Airline_binary_predictions), as.factor(airline_test$satisfaction), positive = "1")
```

## Logistic Model Interpretations 
```{r}

```

#ANN Model
```{r cache=TRUE}
airline_neural<-neuralnet(satisfaction ~., data = airline_train, hidden = 1)
``` 

##ANN Model Predictions 
```{r}
#Won't run with the code using a node 
neural_results<-compute(airline_neural, airline_test[1:ncol(airline_test)])

ANN_predicted<-ifelse(neural_results$net.result>=0.5,1,0)

cor(ANN_predicted, airline_test$satisfaction)

CrossTable(airline_test$satisfaction, ANN_predicted)

#SAME ERROR AS THE ABOVE ONE

#onfusionMatrix(data = as.factor(ANN_predicted), reference = as.factor(airline_test_labels), positive = "1")
```

##ANN Model Optimal 
```{r}

neural_results_2<-predict(airline_neural, newdata=airline_test, type="response")

optimalANN<-optimalCutoff(airline_test$satisfaction, neural_results_2)[1]

neural_airline_binarypred<-ifelse(neural_results_2>optimalANN,1,0)

cor(neural_airline_binarypred, airline_test$satisfaction)

CrossTable(airline_test$satisfaction, neural_airline_binarypred)

#SAME ERROR AS THE ABOVE ONES

#confusionMatrix(data = as.factor(neural_airline_binarypred), reference = as.factor(airline_test_labels), positive = "1")
```

## ANN Model Interpretation 
```{r}

```


#KNN Model
```{r cache=TRUE}

airline_train_KNN <- airline_train
airline_test_KNN <-airline_test

airline_train_KNN$satisfaction<-NULL
airline_test_KNN$satisfaction<-NULL

sqrt(nrow(airline_train_KNN))

str(airline_train_KNN)

summary(airline_train_KNN)
summary(airline_test_KNN)
summary(airline_train_labels)

airline_KNN <- knn(train = airline_train_KNN, test = airline_test_KNN,
                      cl = airline_train_labels, k=325)



CrossTable(x = airline_test_labels, y =airline_KNN, 
           prop.chisq=FALSE)

#SAME ERROR AS ABOVE

#confusionMatrix(as.factor(airline_KNN), as.factor(airline_test_labels), positive = "1")

```

##KNN Model Interpreation 
```{r}

```


#Tree Model 
```{r cache=TRUE}
Airline_Tree<-C5.0(as.factor(satisfaction) ~., data = airline_train)

summary(Airline_Tree)
```

##Tree Model Predictions 
```{r}
airline_tree_pred<-predict(Airline_Tree, airline_test)


CrossTable(airline_test$satisfaction, airline_tree_pred)

#SAME ERROR

#confusionMatrix(data = as.factor(airline_tree_pred), reference = as.factor(airline_test$satisfaction), positive = "1")
```
## Tree Model Interpretation 
```{r}

```

