---
title: 'Project #1'
author: "Kenneth Andrysiak, Joel Revo, Alec Palo, Jacob Cohen, Michael McLaughlan"
date: "2/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data Import and Basic Sturcture Visualization

```{r}
hotelbookings<-read.csv("hotel_bookings.csv")
hotel<-hotelbookings
```

##Data Cleaning

```{r}
hotel$is_canceled<-as.factor(hotel$is_canceled)
hotel$arrival_date_year<-as.factor(hotel$arrival_date_year)
hotel$arrival_date_month<-as.factor(hotel$arrival_date_month)
hotel$arrival_date_day_of_month<-as.factor(hotel$arrival_date_day_of_month)
hotel$arrival_date_week_number<-as.factor(hotel$arrival_date_week_number)
hotel$is_repeated_guest<-as.factor(hotel$is_repeated_guest)
hotel$is_canceled<-as.factor(hotel$is_canceled)
hotel$meal<-as.factor(hotel$meal)
hotel$country<-as.factor(hotel$meal)
hotel$market_segment<-as.factor(hotel$market_segment)
hotel$distribution_channel<-as.factor(hotel$distribution_channel)
hotel$reserved_room_type<-as.factor(hotel$reserved_room_type)
hotel$assigned_room_type<-as.factor(hotel$assigned_room_type)
hotel$deposit_type<-as.factor(hotel$deposit_type)
hotel$agent<-as.factor(hotel$agent)
hotel$company<-as.factor(hotel$company)
hotel$customer_type<-as.factor(hotel$customer_type)
hotel$reservation_status<-as.factor(hotel$reservation_status)
hotel$reservation_status_date<-as.factor(hotel$reservation_status_date)
View(hotel)
summary(hotel)
str(hotel)
```

Here, we cleaned the data to turn a number of different variables in the data set into factors to make the data easier to work with. 

##Analysis and Plotting of Data for VAriables 1 Through 6 in the Hotel Data Set

Here, we will analyze and visualize the data of the first 6 variables listed in the data set provided. We will do both basic and complicated analysis that will lead to the visualization of the data in a number of different plots
