---
title: 'Project #1'
author: "Kenneth Andrysiak, Joel Revo, Alec Palo, Jacob Cohen, Michael McLaughlan"
date: "2/10/2021"
output: 
  html_document: 
    code_folding: hide
    number_sections: true 
    toc: true
    toc_float: 
      collapsed: true  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import and Basic Sturcture Visualization

```{r}
hotel<-read.csv("hotel_bookings.csv")
library(ggplot2)
```

## Data Cleaning

```{r}
hotel$arrival_date_year<-as.factor(hotel$arrival_date_year)
hotel$arrival_date_month<-as.factor(hotel$arrival_date_month)
hotel$arrival_date_day_of_month<-as.factor(hotel$arrival_date_day_of_month)
hotel$arrival_date_week_number<-as.factor(hotel$arrival_date_week_number)
hotel$is_repeated_guest<-as.factor(hotel$is_repeated_guest)
```

```{r}
hotel$meal<-as.factor(hotel$meal)
hotel$country<-as.factor(hotel$country)
hotel$market_segment<-as.factor(hotel$market_segment)
hotel$distribution_channel<-as.factor(hotel$distribution_channel)
hotel$reserved_room_type<-as.factor(hotel$reserved_room_type)
hotel$assigned_room_type<-as.factor(hotel$assigned_room_type)
hotel$deposit_type<-as.factor(hotel$deposit_type)
hotel$agent<-as.factor(hotel$agent)
hotel$company<-as.factor(hotel$company)
hotel$customer_type<-as.factor(hotel$customer_type)
hotel$reservation_status<-as.factor(hotel$reservation_status)
hotel$reservation_status_date<-as.factor(hotel$reservation_status_date)
```

# Introduction 
A number of different variables in the data seem to have relevant business takeaways. In this section, we will analyze these relationships and provide visualizations of the relationships between the variables.

To begin, we will do some basic descriptive analytics on a number of different variables that we think have some pretty significant business insights.

```{r}
summary(hotel$adr)
summary(hotel$required_car_parking_spaces)
summary(hotel$total_of_special_requests)
summary(hotel$reservation_status)
```

*adr*: average daily rate as defined by dividing the sum of all lodging transactions by the total number of staying nights
*required_car_parking_spaces*: number of car parking spaces required by the customer
*total_of_special_requests*: number of special requests made by the customer (e.g. twin bed or high floor)
*reservation_status*: reservation last status, assuming one of three categories: Canceled – booking was canceled by the customer; Check-Out – customer has checked in but already departed; No-Show – customer did not check-in and did inform the hotel of the reason why
*reservation_status_date*: date at which the last status was set. This variable can be used in conjunction with the ReservationStatus to understand when was the booking canceled or when did the customer checked-out of the hotel

## Basic Data Visualization
```{r}
summary(hotel)
str(hotel)
```

Here, we cleaned the data to turn a number of different variables in the data set into factors to make the data easier to work with. 

# Cancelation 

We are focusing on cancellations, because they, logically, have an impact on revenue. We will analyze its connection to other variables in the data we were provided and answer questions that will give insight into how the business should operate moving forward.

## What proportion of reservations are being canceled?
```{r}
tableis_canceled<-table(hotel$is_canceled)
round(prop.table(tableis_canceled),3)
```
What we see here is that roughly 63% of all of the reservations weren't canceled and the other 37% were. What does that mean for us? Well, it gives us context on how many bookings we can list to the public to maximize our revenue with the expectation that people will cancel their reservations. This will allow us to work at optimal capacity and improve efficiency.

## How many of each hotel classification (Resort and City) is being canceled, and at what proportion?
### Number of Resort Hotels Canceled
```{r}
resort<-hotel[hotel$hotel == "Resort Hotel",]
city<-hotel[hotel$hotel == "City Hotel",]
resortcanceled<-hotel[hotel$hotel == "Resort Hotel" & hotel$is_canceled == 1,]
cityiscanceled<-hotel[hotel$hotel == "City Hotel" & hotel$is_canceled == 1,]
nrow(resortcanceled)
```

### Number of City Hotels Canceled
```{r}
nrow(cityiscanceled)
```

### Proportion of Resort Hotels Canceled

```{r}
round(nrow(resortcanceled)/nrow(resort),3)
```

### Proportion of City Hotels Canceled
```{r}
round(nrow(cityiscanceled)/nrow(city),3)
```

### Number of Reservations Canceled by Hotel Classification
```{r}
ggplot(hotel, aes(x=is_canceled, fill = hotel)) + geom_bar() +scale_x_continuous(breaks = seq(0,1))  
```

This graph illustrates the idea presented above in our comments on the cancellation rate. It gives us more context on how many room reservations we should be listing at either hotel classification. We see roughly 41.7% of City Hotels are canceled, compared to 27.8% in Resort Hotels, which tells us that we should be listing more bookings at City hotels to drive revenue.

## Which country of origin had the highest cancelation rate?
```{r}
canceled<-hotel[hotel$is_canceled == 1,]
countrycanceled<-aggregate(hotel$is_canceled, list(Country=hotel$country), mean)
countrycanceled
```
The purpose of answering this question was to get a sense of the different markets that we should be focusing on moving forward, while also learning how to target those markets strategically. 

## What was the average price paid for reservations that were canceled compared to those that weren't? 
```{r}
avgadris_canceled<-tapply(hotel$adr, hotel$is_canceled, mean)
avgadris_canceled
```

## What was the average price paid by those in resort hotels, and those in city hotels? Did this have any effect on the cancellation rate of each of the classifications?
```{r}
avgadrhotel<-tapply(hotel$adr, hotel$hotel, mean)
avgadrhotel
summary(avgadrhotel)
ggplot(hotel, aes(x=adr)) +geom_histogram(binwidth = 25) +xlim(0,350)
```

This is a histogram and statistical data detailing the distribution of price paid by our customers. Our conclusion, based on this graph, is that we should possibly increasing the price of our listings at resort hotels to smooth out the demand distribution and increase out revenue. The fact that the average price paid by our customers is higher than that of our resort hotel average means that we may be able to squeak out more revenue by raising the price of our listings at those resort hotels. 

## Which booking channel had the most cancellations, and by hotel classification?
```{r}
aggregate(hotel$is_canceled, list(hotel=hotel$hotel, DistributionChannel=hotel$distribution_channel), mean)
```

We should evaluate our partnerships with some our distribution channels. More specifically, that should be our partnerships with travel agents and tour operators. They make up the bulk of our reservations, and thus have a large impact on our revenue. If we can increase the quality of these partnerships, we may be able to bring down the cancellation numbers provided in this table. Although our corporate partners also have a relatively large cancellation rate, we recommend that we continue our relationship with them, because they make up a comparatively low volume of our sales.

## How many previous cancellations repeated, and by hotel classification?
```{r}
canceledagain<-ifelse(hotel$is_canceled == hotel$previous_cancellations, 1, 0)
sum(canceledagain)
notcanceledagain<-ifelse(hotel$is_canceled != hotel$previous_bookings_not_canceled, 1, 0)
sum(notcanceledagain)
avgadragain<-tapply(hotel$adr,canceledagain, mean, na.rm = TRUE)
avgadrnot<-tapply(hotel$adr, notcanceledagain, mean, na.rm = TRUE)
avgadragain
avgadrnot
```

This data suggests that we should possibly lean into different markets strategically, as the majority of customers that previously canceled had canceled again. It also appears they pay lower prices compared to those that had not previously canceled with us before. We need to increase the loyalty to our brand, which will increase our retention rate of customers, thus leading to higher cash inflows.

## What was the average number of changes made to the bookings in those bookings that were canceled? What about those that weren't, and by hotel classification?
```{r}
avgchangecanceled<-tapply(hotel$is_canceled, hotel$booking_changes, mean, na.rm = TRUE)
avgchangecanceled
avgchangenot<-tapply(ifelse(hotel$is_canceled == 1, 0,1), hotel$booking_changes, mean, na.rm = TRUE)
avgchangenot
```
It doesn't seem as if the number of changes in the the booking had much effect on the cancellations, however, it would save us cost to decrease the number of changes made to bookings in those reservations that are canceled. At the moment, the bulk of canceled reservations make little changes, which is a good thing. We should continue to look into strategic business plans as a way of decreasing cost in this aspect.

## Did the deposit type have any effect on the chance of cancelation? By hotel classification?
```{r}
deposittypecanceled<-aggregate(hotel$is_canceled, list(DepositType=hotel$deposit_type, Hoteltype=hotel$hotel), mean)
deposittypecanceled
```

Strangely enough, we found that non-refundable deposit types had a grossly large cancellation rate. We should consider changing our listing policy in city hotels to allow for customers not to make a deposit, and allow our resort-goers the opportunity to have a refundable deposit. Our customer segment is a huge part of this, as we believe that the data suggests recreational customers are more inclined to not cancel when they have the ability to get a refund on their deposit, whereas our more "corporate" and city customers are more inclined to not cancel when they don't have to make a deposit in the first place.

# Group Size, Length of Stay, and Type
In this section we will be analyzing the impact of size of group, type of group, and length of stay on different variable such as adr, and number of speical requests. Since adr is directly related to the amount of revenue we will recieve, it is the key variable analyzed in this section. However, we also look at number of special requests, since it is likely that a higher number of special requests impacts the costs for the hotel, thus making a customer less profitable. 

## Group Stay Duration 
```{r}
totalnights<-hotel$stays_in_week_nights+hotel$stays_in_weekend_nights

totalnights <- as.factor(totalnights)

plot(totalnights,ylab="Number of Reservations", xlab="Number of Nights Stayed", main="Number of nights Stayed per Reservation", col="purple")

round(mean(hotel$stays_in_week_nights)+mean(hotel$stays_in_weekend_nights),2)
median(hotel$stays_in_week_nights+hotel$stays_in_weekend_nights)
```
Here we can see the average number of nights stayed by hotel guests. First, we combine stays in weeknights with stays during the weekend to obtain total nights per visit. The plot shows that a majority of guests are staying between 1 to 5 days, with the average being 3.43 nights, and the median being 3 nights. This is substantially higher than the national average stay of about 1.8 days (Business Travel News)
## Stay Duration vs. ADR
```{r}
staydurationadr<-aggregate(adr ~ totalnights, data=hotel, mean, na.rm=TRUE)

staydurationadr

```
Here we can see that add total nights increase, so does ADR, at least minimally, and up to about 6 nights. 
## Group Size
Here, we are looking to find data related to group size. Since the current data has people grouped into three different columns (adults, children and babies) we will first create a variable to look at the total number of people per group. From this, we will look to compare group size to other vairables. 
```{r}


groupsize <-hotel$adults+hotel$children+hotel$babies

groupsize <- as.factor(groupsize)

plot(groupsize, ylab="Number of Reservations", xlab="People in Reservation", main="Group Size per Reservation",col="purple")

mean(hotel$adults+hotel$children+hotel$babies, na.rm = TRUE)
median(hotel$adults+hotel$children+hotel$babies, na.rm = TRUE)


```
Here we combine the adults, children, and babies to find the total group size. On average, reservations contain slightly fewer than 2 people. 


## Group Size Impact on ADR
```{r}
groupsizeadr<-aggregate(adr ~ groupsize, data=hotel, mean, na.rm=TRUE)

groupsizeadr

```
More importantly than just group size, we analyzed the impact of group size on ADR. Here we can see adr increase with group size up to 5 people. Although we don't have data on number of rooms reserved (if more than one) We can see that once the group size is greater than two, revenue per night increases substantially. If in the same number of rooms, the costs also likely do not increase much with added people. 

## Market Segment Group Type vs. ADR
Here were looking at groups by market segment. 
```{r}
library(ggplot2)

ggplot(hotel, aes(x=groupsize, fill=market_segment, main= "Number of Reservations by Group size and Market Segment")) + geom_bar(position = "stack")

marketsegmentadr<-aggregate(adr ~ market_segment, data=hotel, mean, na.rm=TRUE)


plot(hotel$market_segment, main="Number of Reservations by Market Segment", las=2, col="purple")



marketsegmentadr

```
From the above graphs and information, we can see that group size doesn't have a noticiable impact on the market segment that books a reservation. More interestingly though, is looking at the data regarding which market segments book most frequently, and the average ADR of those segments. Online TA and Offline TA TO are the most frequently used booking methods, however, Online is more than $30 higher per day in ADR than offline booking (see table). This might suggest that the Hotel should shift towards exclusively online booking, as it is possible the room disocount are higher for offline booking. 

## Kids Impact on Special Requests 
Here, we will look to see if number of children + babies is related to the number of special requests, as we suspect that a higher number of special requests will lead to larger costs for the hotel.  
```{r}

mean(hotel$total_of_special_requests, na.rm=TRUE)

groupswithoutkids<-hotel[hotel$babies ==0 & hotel$children ==0,] #Groups with no kids
groupswithkids <- hotel[hotel$babies> 0 | hotel$children> 0,] #Groups with at least 1 child or baby

mean(groupswithkids$total_of_special_requests, na.rm=TRUE)

mean(groupswithoutkids$total_of_special_requests, na.rm = TRUE)




```

```{r}
hotel$TotalKids <- hotel$babies + hotel$children
lm_kidsSpecReq <- lm(total_of_special_requests ~ TotalKids, data = hotel)
summary(lm_kidsSpecReq)

#ggplot(data = hotel) + geom_point(mapping = aes(TotalKids, total_of_special_requests))
ggplot(data = hotel) + geom_bar(mapping = aes(TotalKids, main = "Bar Plot of Total Number of Kids"))
```

Here we can see very interesting results that confirm our hypothesis. First, the average number of special requests for all reservations is .57. However, groups with children or babies almost average double that number at .927 per reservation while groups without any children or babies average only .54 special requests.  We can also see from the linear model that there is in fact correlation between the number of special requests and the amount of total kids in a reservation. 

## Kids Impact on Special Requests and ADR
```{r}
groupswithkids <- hotel[hotel$adults>=2 & hotel$babies> 0 | hotel$children> 0, ]

groupswithadults <- hotel[hotel$adults>=3,]

mean(groupswithadults$adr)

mean(groupswithkids$adr, na.rm=TRUE)
```
In addition, we wanted to check if groups with kids are spending more money on average (to potentially cancelout the cost of more special requests). However, from these results we can see that the adr of groups with children and babies and the adr of groups with just adults is roughly the same. This means that groups with kids are much more costly to the hotel without driving additional revenue. 

# Minimizing Costs
One of the most damaging things to a hotel's profit is special request accommodation. This section will address the variables that impact the total number of special requests, including time of year and number of children. Additionally, the section takes a look at the number of guests requiring parking spaces, in order to draw insights that could help hotels allocate their space most effectively.

## Average Price Per Person 

***What is this code chunk intended to do?
```{r}
priceperperson <- hotel$adr*groupsize
```

## Histogram of Special Requests
```{r}
ggplot(data = hotel) + geom_histogram(mapping = aes(total_of_special_requests), fill = "purple")
table(hotel$total_of_special_requests)
#nrow(hotel$total_of_special_requests[hotel$total_of_special_requests == "0"])
#nrow(hotel[hotel$total_of_special_requests == 0])
```
While the majority of hotel guests have 0 special requests, there is still a significant of guests with at least 1 special request. Considering the potential costs that special requests can impose, it is important to dive deeper into what causes them.

## Avg Daily Rate by Type of Hotel
```{r}
mean(hotel$adr[hotel$hotel == "Resort Hotel"]) # avg daily rate for resort hotel
mean(hotel$adr[hotel$hotel == "City Hotel"]) # avg daily rate for city hotel
```
On average, the average daily rate at city hotels is greater than the average daily rate at resort hotels.

## Required Parking Spaces by Hotel Type
```{r}
ZeroParkingSpaces <- hotel[hotel$required_car_parking_spaces == "0",] # Filter for guests with 0 parking spaces
nrow(ZeroParkingSpaces) # Print total number of guests with 0 parking spaces)

SomeParkingSpaces <- hotel[hotel$required_car_parking_spaces > "0",] # Filter for guests with 1+ parking spaces
nrow(SomeParkingSpaces) # Print total number of guests with 1+ parking spaces)

ZeroParkingSpacesCity <- hotel[hotel$required_car_parking_spaces == "0" & hotel$hotel == "City Hotel",] # Filter for guests at city hotel requiring 1+ parking spaces
SomeParkingSpacesCity <- hotel[hotel$required_car_parking_spaces > "0" & hotel$hotel == "City Hotel",] # Filter for guests at city hotel requiring 1+ parking spaces

ZeroParkingSpacesResort <- hotel[hotel$required_car_parking_spaces == "0" & hotel$hotel == "Resort Hotel",] # Filter for guests at resort hotel requiring 1+ parking spaces
SomeParkingSpacesResort <- hotel[hotel$required_car_parking_spaces > "0" & hotel$hotel == "Resort Hotel",] # Filter for guests at resort hotel requiring 1+ parking spaces

nrow(ZeroParkingSpacesCity) # Number of guests at city hotels requiring zero parking spaces
nrow(SomeParkingSpacesCity) # Number of guests at city hotels requiring 1+ parking spaces

nrow(ZeroParkingSpacesResort) # Number of guests at resort hotels requiring zero parking spaces
nrow(SomeParkingSpacesResort) # Number of guests at resort hotels requiring 1+ parking spaces

nrow(ZeroParkingSpacesCity) / nrow(hotel) # Proportion of guests at city hotels that do not require parking spaces
nrow(ZeroParkingSpacesResort) / nrow(hotel) # Proportion of guests at resort hotels that do not require parking spaces

ggplot(data=hotel) + geom_bar(mapping = aes(required_car_parking_spaces, fill = hotel)) # visual of parking spaces by hotel type
```
Interestingly, approximately 64% of guests at city hotels do not require parking spaces, while only approximately 29% of guests at resort hotels do not require parking spaces. In other words, guests at resort hotels at much more likely to need a parking space than at city hotels. This can help hotels in terms of allocating space; resort hotels need to designate more of their space for parking, while city hotels can afford to use some of their space for other purposes.


## Does the time of year influence the number of special requests?

```{r}
MonthsTable <- table(hotel$arrival_date_month, hotel$total_of_special_requests)
MonthsTable
#How to order months
#How to pull insights
```



```{r}
SpecReqByMonth <- aggregate(hotel$total_of_special_requests, list(Month=hotel$arrival_date_month), mean) # Avg SpecReq by month
SpecReqByMonth
round(sort(SpecReqByMonth[,2], decreasing = T),2) # Need help keeping the month labels when I sort

ggplot(data = hotel) + geom_bar(mapping = aes(arrival_date_month, fill = "red"))
```
On average, a guest is more likely to ask for a special request(s) in the summer months (July and August). On the contrary, they are less likely to ask for a special request(s) in the colder months (January, February, March, September, and October)

## Is there a linear relationship between avg daily rate and special requests

```{r}
lm_SpecReq_ADR <- lm(adr ~ total_of_special_requests, data=hotel) # create linear model b/w special requests and adr
summary(lm_SpecReq_ADR)
```
This linear model tests the predictive power of total number of special requests on average daily rate. Since the p-value of the model is so low, we can claim that number of special requests is a significant predictor of average daily rate. On average, for every additional special request made, the average daily rate of that guest increases by about $11.

## Relationship between reservation status and demographics of guests (aduls/children/babies, country)

```{r}
Table_ResStatus_Kids <- table(hotel$reservation_status, hotel$TotalKids)
Table_ResStatus_Kids
# calculate proportions
```
Regardless of the total number of kids, the most likely reservation status is "Check-Out," meaning that guests tend to follow through with their hotel stays. Additionally, it seems that the more kids a party has, the less likely they are to cancel or not show up.

# Hotel Room Reservations and Market Segment Analysis

## What are the reservation requests for each room?

```{r}
hotel$reserved_room_type <- as.factor(hotel$reserved_room_type)

table(hotel$reserved_room_type)

#This table and barplot shows how room A is the most reserved room, followed by room D and E. 

reserved_room_type <- sort(table(hotel$reserved_room_type))

barplot(reserved_room_type, main = "# Room requests", xlab = "month", ylab = "# of room requests", cex.names = 0.4, col = "blue")

#See above for a visual representation of room reservation requests. 

data.frame(reserved_room_type)

##Which hotel rooms are requested the most by market segment?

library(dplyr)

#count(hotel, market_segment, reserved_room_type)

library(ggplot2)
 
# Stacked

hotel$count <- 1

ggplot(hotel, aes(fill=reserved_room_type, y=count, x= market_segment)) + geom_bar(position ="fill", stat="identity")

#The graph above shows the mix of room bookings for every market segment. This information is useful to predict the demand for rooms by each customer market segment.

```


##What insights can we draw from the market segment of our clients and how it relates to repeat customers?

We have discovered that the majority of repeat customers are employed in the aviation, corporate or complementary market segments. 

```{r}
aggregate(is_repeated_guest ~ market_segment, data = hotel, mean)

market_segment <- sort(table(hotel$market_segment))

barplot(market_segment, main = "Market Segment", xlab = "Market Segment", ylab = "Frequency", cex.names = 0.4, col = "blue")

#Online TA is the most popular method of booking a reservation with a hotel. 

ggplot(hotel, aes(x=market_segment, y=previous_cancellations, fill=reserved_room_type)) + 
  geom_bar(stat = "identity")

```

##Cancellation Breakdown

37.04% of reservations are eventually canceled by the hotel's customers. 

```{r}
nrow(hotel[hotel$is_canceled == 1,]) / nrow(hotel)  

#Total canceled reservations out of all entries (37.04%)

num_cancelled_city <- nrow(hotel[hotel$is_canceled == 1 & hotel$hotel == "City Hotel",])

print(num_cancelled_city)

num_city_bookings <- nrow(hotel[hotel$hotel == "City Hotel",])

print(num_city_bookings)

print(num_cancelled_city/num_city_bookings)  #Proportion canceled in City Hotel (41.7%)
```


##Is there a connection between the market segment of our clients and their liklihood of being a repeat customer?

```{r}
hotel$market_segment <- as.factor(hotel$market_segment)

hotel$distribution_channel <- as.factor(hotel$distribution_channel)

aggregate(is_repeated_guest ~ market_segment, data = hotel, mean)

##Corporate distribution is the largest source of repeat business for the hotel industry.

aggregate(is_repeated_guest ~ distribution_channel, data = hotel, mean)


```

##How many guests recieved the hotel room they requested?

```{r}
hotel$reserved_room_type <- as.factor(hotel$reserved_room_type)

hotel$reserved_room_type<- factor(hotel$reserved_room_type, levels = c(levels(hotel$reserved_room_type ),"I", "K"))

hotel$assigned_room_type <- as.factor(hotel$assigned_room_type)

hotel$Room_Match <- ifelse(hotel$reserved_room_type == hotel$assigned_room_type, 1, 0)

Room_Match <- as.numeric(hotel$Room_Match)

Percent_Room_Match <- sum(hotel$Room_Match)/nrow(hotel)

Percent_Room_Match

levels(hotel$assigned_room_type)

levels(hotel$reserved_room_type)


```


##Previous Cancellations & Previous Bookings not Canceled 

```{r}

library(ggplot2)

aggregate(previous_cancellations ~ market_segment, data = hotel, mean)

##Corporate accounts for the highest number of bookings not canceled. 

hotel$total <- "total"

ggplot(hotel, aes(fill=market_segment, y=previous_cancellations, x= total)) + geom_bar(position ="fill", stat="identity")

##The graph above shows the breakdown of previous cancellations by market segment. The hotel can use this data to estimate the number of rooms that will be cancelled according to market segment hotel booking data. 

#The table below shows the total number of previous cancellations by market segment. 

aggregate(previous_bookings_not_canceled ~ market_segment, data = hotel, sum)
```


# Cancellations, Deposit Type, Customer Type
[description of section]

## Overall cancellation rate

```{r}
cancelration_rate <- (sum(hotel$is_canceled))/nrow(hotel)
round(cancelration_rate,2)
```


## Adr by customer type

In this section we will be looking at the average daily rate for each customer type - contract, group, transient and transient-party. This is important as knowing which customer group has the highest average daily rate will allow this hotel to market to that group, bringing more revenue.

```{r}
adr_by_ctype <- tapply(hotel$adr, hotel$customer_type, mean, na.rm = T) # This function creates a table showing the average daily rates for each customer type

adr_by_ctype <- data.frame(adr_by_ctype) # This function transforms the table into a data frame, allowing it to be used in a ggplot

adr_by_ctype <- cbind(newColName = rownames(adr_by_ctype), adr_by_ctype) 
rownames(adr_by_ctype) <- 1:nrow(adr_by_ctype) # These two functions give the first collumn of the dataframe a name, so it can be used in ggplot

names(adr_by_ctype)[names(adr_by_ctype) == "newColName"] <- "customer"
names(adr_by_ctype)[names(adr_by_ctype) == "adr_by_ctype"] <- "adr" # This group of functions changes the names of the collums to better reflect what they are, customer type and adr

adr_by_ctype # This is the final data frame that will be graphed below

ggplot(adr_by_ctype, aes(x = customer, y = adr)) + geom_bar(stat = "identity", fill = "#FF9999", color = "black") + ggtitle("Average Daily Rate by Customer Type") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x = "Customer Type", y = "Average Daily Rate", title = "Average Daily Rate by Customer Type") # This bar graph shows the average daily rates for each customer type. It is a bar graph as it is easy to see the differences between them
```
From this data we can see that transient parties, on average, have the highest average daily rate. This means that hotels should target transient parties, which would allow them to increase revenue.

## Cancelations by deposit type

This section will explore the number of cancellations for each deposit type. By understanding which deposit type has the most cancellations you can attempt to limit the number of cancellations or book rooms accordingly.

```{r}
canceled_by_deposite <- table(hotel$is_canceled, hotel$deposit_type) # This function creates a table of cancellations by deposite type, 1 being canceled, 0 being not canceled

canceled_by_deposite <- data.frame(canceled_by_deposite) # this function turns that table into a dataframe
canceled_by_deposite

ggplot(canceled_by_deposite, aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(position = "dodge", stat = "identity", color = "black") + ggtitle("Cancelations by Deposite Type") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x = "Not Canceled / Canceled", y = "Number of Observations", title = "Cancelations by Deposite Type") + guides(fill=guide_legend(title = "Deposite Type")) # This bar graph shows canceled and not canceled for the different deposite types

```

## Cancelations by Customer type

Being able to see how many cancelations by each customer type will allow hotels to properly book - overbook based on how likely groups are to cancel - rooms and mitigate number of cancellations. 

```{r}
canceled_by_Ctype <- tapply(hotel$is_canceled, hotel$customer_type, sum, na.rm = T) # This function creates a table of cancelations by customer type

canceled_by_Ctype <- data.frame(canceled_by_Ctype) # This function transforms the previous table into a dataframe

canceled_by_Ctype <- cbind(newColName = rownames(canceled_by_Ctype), canceled_by_Ctype)
rownames(canceled_by_Ctype) <- 1:nrow(canceled_by_Ctype) # This group of functions adds a collumn name allowing it to be graphed in a ggplot

names(canceled_by_Ctype)[names(canceled_by_Ctype) == "newColName"] <- "Ctype"
names(canceled_by_Ctype)[names(canceled_by_Ctype) == "canceled_by_Ctype"] <- "Can" # These two functions change the names of the columns to better describe them

canceled_by_Ctype

ggplot(canceled_by_Ctype, aes(x = Ctype, y = Can)) + geom_bar(stat = "identity", fill = "#FF9999", color = "black") + ggtitle("Cancelations by Customer Type") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x="Customer Type", y="Number of Cancelations", title="Cancelations by Customer Type") # This bar graph shows the number of cancelations for each customer type. It is a bar graph as it is easy to compare different groups cancelations.
```

## Deposites by customer type

Previously we explored how many cancelations by deposite type. Knowing which groups are most likely to use which deposit type will enable hotels to further predict cancelations. This information will further allow hotels to more accuretly predict cancellations and book hotel rooms better.

```{r}
deposite_by_Ctype <- table(hotel$deposit_type, hotel$customer_type) # This function creates a table of deposite types by customer types

deposite_by_Ctype <- data.frame(deposite_by_Ctype)[,] # This function transforms that table into a dataframe, allowing it to be plotted in ggplot

names(deposite_by_Ctype)[names(deposite_by_Ctype) == "Var1"] <- "deposite_type"
names(deposite_by_Ctype)[names(deposite_by_Ctype) == "Var2"] <- "group_type"
names(deposite_by_Ctype)[names(deposite_by_Ctype) == "Freq"] <- "count" # This group of functions changes the names of the collumns to better describe them

deposite_by_Ctype

ggplot(data = deposite_by_Ctype, aes(x = deposite_type, y = count, fill = group_type)) + geom_bar(position = "dodge", stat = "identity", color = "black") + ggtitle("Deposite Type by Customer Type") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x="Deposite Type", y="Number of Customers", title="Deposite Type by Customer Type") + guides(fill=guide_legend(title = "Customer Type")) # A bar graph was used to better visualize the number of customers in each type by the deposite type
```

## Average daily rates by company

Knowing which company has the highest average daily rates will enable higher revenue as they can attempt to increase bookings from that company.

```{r}
mean_adr_by_co <- tapply(hotel$adr, hotel$company, mean, na.rm = T) # This function creates a table showing the average daily rates by each company. 

mean_adr_by_co <- data.frame(mean_adr_by_co) # This function transforms the previous table into a dataframe.

mean_adr_by_co <- cbind(newColName = rownames(mean_adr_by_co), mean_adr_by_co)
rownames(mean_adr_by_co) <- 1:nrow(mean_adr_by_co) # This group of functions adds a name to one of the columns.

names(mean_adr_by_co)[names(mean_adr_by_co) == "newColName"] <- "co"
names(mean_adr_by_co)[names(mean_adr_by_co) == "mean_adr_by_co"] <- "adr" # This group of functions changes the names of the collumns to better describe them.

summary(mean_adr_by_co)

ggplot(mean_adr_by_co, aes(x = co, y = adr)) + geom_point(size = 1, color = "#FF9999") + ggtitle("Average Daily Rate by Company") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x="Company", y="Average Daily Rate", title="Average Daily Rate by Company") # This scatter plot shows the average daily rates for each company.
```

## Average daily rates by agents

Knowing which agents has the highest average daily rates will enable higher revenue as they can attempt to increase bookings from those agents.

```{r}
mean_adr_by_agent <- tapply(hotel$adr, hotel$agent, mean, na.rm = T) # This function creates a table showing the average daily rates by each agent. 


mean_adr_by_agent <- data.frame(mean_adr_by_agent) # This function transforms the previous table into a dataframe.

mean_adr_by_agent <- cbind(newColName = rownames(mean_adr_by_agent), mean_adr_by_agent)
rownames(mean_adr_by_agent) <- 1:nrow(mean_adr_by_agent) # This group of functions adds a name to one of the columns.

names(mean_adr_by_agent)[names(mean_adr_by_agent) == "newColName"] <- "agent"
names(mean_adr_by_agent)[names(mean_adr_by_agent) == "mean_adr_by_agent"] <- "adr" # This group of functions changes the names of the collumns to better describe them.

summary(mean_adr_by_agent)

ggplot(mean_adr_by_agent, aes(x = agent, y = adr)) + geom_point(size = 1, color = "#FF9999") + ggtitle("Average Daily Rate by Agent") + theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5)) + labs(x = "Agent ID", y = "Average Daily Rate ($)", title = "Average Daily Rate by Agent") # This scatter plot shows the average daily rates for each agent.
```

# Time of Year 

## Do Holidays impact ADR? 
```{r}
mean(hotel$adr) #Find Adr on average



tableJuly4 = hotel[hotel$arrival_date_month == "July" & hotel$arrival_date_day_of_month== 3 | hotel$arrival_date_day_of_month==4,] #Create Table to isolate specifc columns that we want

round(mean(tableJuly4$adr, na.rm= TRUE),2) #Find mean for that holiday 


tablenewyearseve = hotel[hotel$arrival_date_month == "December" & hotel$arrival_date_day_of_month== 31,]
round(mean(tablenewyearseve$adr, na.rm= TRUE),2) #repeat

tableXmas = hotel[hotel$arrival_date_month == "December" & hotel$arrival_date_day_of_month== 24 | hotel$arrival_date_day_of_month==25,]
round(mean(tableXmas$adr, na.rm= TRUE),2)

tableValentines = hotel[hotel$arrival_date_month == "February" & hotel$arrival_date_day_of_month== 13 | hotel$arrival_date_day_of_month==14,]
round(mean(tableValentines$adr, na.rm= TRUE),2)

```
From the above information we have pretty interesting findings. When we look at the average ADR across the entire year, and compare that to the ADR for reservations during the Holidays, we see that the Hotel is actually charging the same price or lower for most holidays. Some of the typically most expensive holidays to travel (July 4th, Xmas, Valentines) have roughly the same ADR as the overall average. Only the ADR for New Years is higher, but only by 29%, whereas some hotes raise prices by as much as 100% during busy holidays. 

## Does Resort type impact ADR on holidays?

```{r}
tableResort = hotel[hotel$hotel == "Resort Hotel",] #Here we are doing the same code as above, but also isolation for type of hotel to identify whether this is a hotel specifc problem 
mean(tableResort$adr)

tableJuly4Resort = hotel[hotel$arrival_date_month == "July" & hotel$arrival_date_day_of_month== 3 | hotel$arrival_date_day_of_month==4 & hotel$hotel== "Resort Hotel",]

round(mean(tableJuly4Resort$adr, na.rm= TRUE),2)
tableCity = hotel[hotel$hotel == "City Hotel",]
mean(tableCity$adr)

tableJuly4City = hotel[hotel$arrival_date_month == "July" & hotel$arrival_date_day_of_month== 3 | hotel$arrival_date_day_of_month==4 & hotel$hotel== "City Hotel",]

round(mean(tableJuly4City$adr, na.rm= TRUE),2)


```
```{r}
tableXmasCity = hotel[hotel$arrival_date_month == "December" & hotel$arrival_date_day_of_month== 24 | hotel$arrival_date_day_of_month==25 & hotel$hotel=="City Hotel",]
round(mean(tableXmasCity$adr, na.rm= TRUE),2)


TableXmasResort = hotel[hotel$arrival_date_month == "December" & hotel$arrival_date_day_of_month== 24 | hotel$arrival_date_day_of_month==25 & hotel$hotel=="Resort Hotel",]
round(mean(TableXmasResort$adr, na.rm= TRUE),2)


```
From the above information we can also see that the problem is not hotel specifc. Both City and Resort hotels charge roughly the same ADR overall throughout the year as they do on a few key holidays (shown here are Christmas and July 4th)

# Type of Client 
## Which clients have the highest adr? 
## Which clients have the most special requests?
## Which clients have the highest cancellation rate 

# Summary

In conclusion, after analyzing the hotel data set we can draw the following conclusions and have the following recommendations. 

1: Cancelation Section

2: Group Size, Type and Length of Stay Section 
The largest takeaways from this section are as follows. Attempt to make reservations through Online TA market segment, which yields the highest ADR on average. In addition, it is important to look to minimize special requests, as they are likely tied to higher costs. More kids in groups means more special requests, thus the hotel might want to charge a higher fee if there are many kids in a reservation. 


3. Minimizing Costs

4. Michael Section 

5. Alec Section 

6. Time of Year Section
Here, we discovered that the ADR (for both hotels) is roughly the same for the hotel during holiday and non holiday bookings. Thus, we recomend that the hotel begin to increase rates and prices during the holiday seasons, which is typical for any hotel. 



