---
title: 'Project #1'
author: "Kenneth Andrysiak, Joel Revo, Alec Palo, Jacob Cohen, Michael McLaughlan"
date: "2/10/2021"
output: 
  html_document: 
    code_folding: hide
    number_sections: true 
    toc: true
    toc_float: 
      collapsed: true  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Data Import and Basic Sturcture Visualization

```{r}
hotelbookings<-read.csv("hotel_bookings.csv")
hotel<-hotelbookings
```

#Data Cleaning

```{r}
hotel$hotel<-as.factor(hotel$hotel)
hotel$is_canceled<-as.factor(hotel$is_canceled)
hotel$arrival_date_year<-as.factor(hotel$arrival_date_year)
hotel$arrival_date_month<-as.factor(hotel$arrival_date_month)
hotel$arrival_date_day_of_month<-as.factor(hotel$arrival_date_day_of_month)
hotel$arrival_date_week_number<-as.factor(hotel$arrival_date_week_number)
hotel$is_repeated_guest<-as.factor(hotel$is_repeated_guest)
hotel$is_canceled<-as.factor(hotel$is_canceled)
```

```{r}
hotel$meal<-as.factor(hotel$meal)
hotel$country<-as.factor(hotel$meal)
hotel$market_segment<-as.factor(hotel$market_segment)
hotel$distribution_channel<-as.factor(hotel$distribution_channel)
hotel$reserved_room_type<-as.factor(hotel$reserved_room_type)
hotel$assigned_room_type<-as.factor(hotel$assigned_room_type)
hotel$deposit_type<-as.factor(hotel$deposit_type)
hotel$agent<-as.factor(hotel$agent)
hotel$company<-as.factor(hotel$company)
hotel$customer_type<-as.factor(hotel$customer_type)
hotel$reservation_status<-as.factor(hotel$reservation_status)
hotel$reservation_status_date<-as.factor(hotel$reservation_status_date)
```

## Basic Data Visualization
```{r}
summary(hotel)
str(hotel)
```

Here, we cleaned the data to turn a number of different variables in the data set into factors to make the data easier to work with. 

#Possible Relationships in the Data

A number of different variables in the data seem to have relevant business takeaways. In this section, we will analyze these relationships and provide visualizations of the relationships between the variables.

To begin, we will do some basic descriptive analytics on a number of different variables that we think have some pretty significant business insights.


```{r}
tableis_canceled<-table(hotel$is_canceled)
round(prop.table(tableis_canceled),3)

```

##How big are the groups that are staying, and how long are they staying?
```{r}

adults <-hotel$adults
children <-hotel$children
babies <- hotel$babies
totalpeople <-adults+children+babies

totalpeople <-as.factor(totalpeople)

plot(totalpeople,ylab="Number of Reservations", xlab="People in Reservation")


weekendnights <- hotel$stays_in_weekend_nights
weekdaynights <- hotel$stays_in_week_nights
totalnights <-weekendnights+weekdaynights

totalnights <- as.factor(totalnights)

plot(totalnights,ylab="Number of Reservations", xlab="Number of Nights Stayed")




```


#Summary


