---
title: 'Project #2'
author: "Kenneth Andrysiak, Joel Revo, Alec Palo, Jacob Cohen, Michael McLaughlan"
date: "2/10/2021"
output: 
  html_document: 
    code_folding: hide
    number_sections: true 
    toc: true
    toc_float: 
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction 

The purpose of this report is threefold. First, we want to identify what predictor variables lead to borrowers on prosper obtaining a higher interest rate. This will give us an idea of what the market thinks are "credit risks." Second, we will make a logistic model to determine what predictor variables are leading to atual loan defaults. From there, we will then be able to compare these two models, to see if the same variables leading to higher interest rates are also the ones causing defaults. If there is a difference between the two, there is likely an arbitrage opportunity. This could mean the market thinks one predictor is a risk that actually isn't, or, a risk that isn't being considered by the market is in fact causing defaults. 

# Data Import and Basic Structure Visualization
```{r}
listings<-read.csv("ProjectA_Listings2013.csv")
str(listings)
listings<-na.omit(listings)
listingsGLM <- listings
```

# Data Cleaning For Linear Regression
```{r}
listings$number_of_days<-NULL
listings$principal_balance<-NULL
listings$loan_status<-NULL
listings$loan_status_description<-NULL
listings$loan_origination_date<-NULL
listings$listing_category_id<-NULL
listings$prosper_rating<-as.factor(listings$prosper_rating)
listings$listing_term<-NULL
listings$listing_monthly_payment<-NULL
listings$scorex<-as.factor(listings$scorex)
listings$income_range<-NULL
listings$income_range_description<-as.factor(listings$income_range_description)
listings$income_verifiable<-as.factor(listings$income_verifiable)
listings$dti_wprosper_loan<-NULL
listings$employment_status_description<-as.factor(listings$employment_status_description)
listings$occupation<-as.factor(listings$occupation)
listings$borrower_state<-NULL
listings$borrower_city<-NULL
listings$delinquencies_last7_years<-NULL
listings$lender_indicator<-NULL
listings$first_recorded_credit_line<-NULL
listings$public_records_last10_years<-NULL
listings$was_delinquent_derog<-NULL
```

# Linear Regression Model
```{r}
lr_model_1<-lm(borrower_rate~.,data = listings)

step_model_1<-step(lr_model_1)
summary(step_model_1)
```

## Linear Regression Intrepretation:

The significant Variables for the Linear Regression Model includes:

- loan_status_description 
- prosper_rating
- amount_funded
- scorex
- stated_monthly_income
- employment status description
- occupation
- current_delinquencies
- public_records_last12_months
- credit_lines_last7_years
- inquiries_last6_months
- amount delinquent
- current_credit_lines
- total inquiries
- satisfactory_accounts
- delinquencies_over30_days

These variables are significant in determining how loans are priced, or how the interest rate for a loan is determined. Therefore the market considers these factors credit risks, and below we will analyze the impact and rationale behind each factor. 

The loan status description is only significant on two levels, including when a loan is current and completed. When an individual has a loan currently or has completed a loan in the past their interest rate is reduced, likely from their perceived credit worthiness as someone already gave them a loan. 

The amount of the loan that is funded was also significant and positively correlated to the interest rate. Our speculation is that individuals who are deemed risky were forced to have a higher amount funded than other individuals and were given a higher rate. Therefore the amount funded is a function of the interest rate, not the other way around. 

Prosper rating is significant on all levels, and scorex is significant with 7/10 of its levels. Individuals with AA prosper rating have their interest rate reduced on average, with all other levels increasing the interest rate. Scorex is positively correlated with the interest rate and as the credit score increases, the addition to the interest rate decreases. This makes sense as someone with a high credit score or prosper ratings is more financially secure. 

Stated monthly income is negatively correlated with the interest rate, which makes sense as those with high incomes are more financially stable and are more likely to pay back a loan.

Employment status was significant on three levels, including full time, other and self employed. Full time workers and 'other' increase your interest rate, which makes sense since other job descriptions such as part time increase the interest rate even more. 

Most of the occupation choices are significant positively correlated to interest rate. Occupations with higher average salaries have their interest rate increased by less than those with a lower salary job. This is not always the case however.

Current delinquencies, amount delinquent and delinquencies over 30 days are unsurprisingly positively correlated to the interest rate and significant, which makes sense as creditors are unlikely to lend to someone who already has unpaid debt. 

Public records last 12 months, credit lines last 7 years, and inquiries last 6 months, total inquiries,and current credit lines are positively correlated with the interest rate, showing that outstanding credit lines and a history of loans increases the interest rate. This may be because the person is more likely to have excess debt to pay back, or is more likely to take on additional debt in the future. All of these factors point to the borrower's credit history. 
Satisfactory accounts is significant and negatively correlated with the interest rate, which makes sense since healthy loans and financial accounts is a positive indication of someone's creditworthiness. 

```{r}
library(class)
library(tidyverse)
library(ggplot2)
library(vctrs)
library(caret)
library(gmodels)
library(e1071)
library(dplyr)
library(speedglm)
library(Matrix)
library(MASS)
library(fastglm)
```

# Data Cleaning Logistic Regression 
```{r}

listingsGLM$prosper_rating<-as.factor(listingsGLM$prosper_rating)
listingsGLM$income_range<-as.factor(listingsGLM$income_range)
listingsGLM$income_range_description<-as.factor(listingsGLM$income_range_description)
listingsGLM$employment_status_description<-as.factor(listingsGLM$employment_status_description)
listingsGLM$scorex<-as.factor(listingsGLM$scorex)
listingsGLM$occupation <- as.factor(listingsGLM$occupation)
listingsGLM$first_recorded_credit_line <- as.Date(listingsGLM$first_recorded_credit_line, format = "%m/%d/%Y")
listingsGLM$loan_origination_date <- as.Date(listingsGLM$loan_origination_date, format = "%m/%d/%Y")
listingsGLM$loan_status_description <- ifelse(listingsGLM$loan_status_description == "CHARGEOFF", 1, ifelse(listingsGLM$loan_status_description == "DEFAULT", 1, ifelse(listingsGLM$loan_status_description == "COMPLETED", 0, ifelse(listingsGLM$loan_status_description == "CURRENT", 0, NA))))


listingsGLM$loan_status <- NULL
listingsGLM$borrower_city <- NULL
listingsGLM$borrower_state <- NULL
#listingsGLM$prosper_rating <- NULL
#listingsGLM$income_range_description <- NULL
#listingsGLM$occupation <- NULL
listingsGLM$total_trade_items <- NULL
listingsGLM$now_delinquent_derog <- NULL
#listingsGLM$delinquencies_over90_days <- NULL
listingsGLM$number_of_days<-NULL
listingsGLM$listing_term<-NULL
listingsGLM$listing_monthly_payment<-NULL
listingsGLM$principal_balance<-NULL
listingsGLM$loan_origination_date<-NULL
```

# Logistic Regression Model

Our first task was to look at a logistic model using the same variables that were significant in our linear regression model. 

```{r}
set.seed(123)
test_set<-sample(1:nrow(listingsGLM), 8385)
listingsGLM_train<-listingsGLM[-test_set,]
listingsGLM_test<-listingsGLM[test_set,]
```

```{r}
logit_group_1<-glm(loan_status_description ~ stated_monthly_income + income_verifiable + amount_funded + prosper_rating + scorex + prosper_score + stated_monthly_income + income_verifiable + employment_status_description + occupation + current_delinquencies + public_records_last12_months + credit_lines_last7_years + inquiries_last6_months + amount_delinquent + current_credit_lines + open_credit_lines + bankcard_utilization + installment_balance + total_inquiries + satisfactory_accounts + delinquencies_over30_days + 
    delinquencies_over60_days ,data=listingsGLM, family = "binomial" )

summary(logit_group_1)

```

After removing the variables that are insignificant, we are left with the following model. 
```{r}
logit_group_2<-glm(loan_status_description ~ stated_monthly_income + income_verifiable + amount_funded + prosper_rating + stated_monthly_income + income_verifiable + credit_lines_last7_years + bankcard_utilization + installment_balance + total_inquiries + satisfactory_accounts + delinquencies_over30_days + 
    delinquencies_over60_days ,data=listingsGLM, family = "binomial" )


summary(logit_group_2)


```

# Logistic Regression Model - Loan Status Description vs. entire data set 

However, in order to make a better model, we decided that it would be better to run "loan_status_description" against every variable in the dataset. Then, one by one, the insignificant variables were removed. 
```{r}
logit_group_3<-glm(loan_status_description ~ .,data=listingsGLM, family = "binomial" )

summary(logit_group_3)
```

Here, we tested the model and slowly removed variables until only significant ones remained. This was done in one step here for the simplicity and to be succinct. (in practice we ran the model after removing each variable, to check to see if other variables became significant)
```{r}

listingsGLM2<-listingsGLM
listingsGLM2$income_range_description <-NULL
listingsGLM2$occupation<-NULL
listingsGLM2$listing_category_id<-NULL
listingsGLM2$months_employed<-NULL
listingsGLM2$delinquencies_last7_years<-NULL
listingsGLM2$was_delinquent_derog<-NULL
listingsGLM2$public_records_last12_months<- NULL
listingsGLM2$delinquencies_over30_days<-NULL
listingsGLM2$total_inquiries<-NULL
listingsGLM2$first_recorded_credit_line<-NULL
listingsGLM2$real_estate_balance<-NULL
listingsGLM2$revolving_balance<-NULL
listingsGLM2$installment_balance<-NULL
listingsGLM2$total_open_revolving_accounts<-NULL
listingsGLM2$real_estate_payment<-NULL
listingsGLM2$bankcard_utilization<-NULL
listingsGLM2$open_credit_lines<-NULL
listingsGLM2$inquiries_last6_months<-NULL
listingsGLM2$dti_wprosper_loan<-NULL
listingsGLM2$income_verifiable<-NULL
listingsGLM2$employment_status_description<-NULL
listingsGLM2$lender_indicator<-NULL
listingsGLM2$current_credit_lines<-NULL
listingsGLM2$installment_balance<-NULL
listingsGLM2$prosper_score<-NULL
listingsGLM2$income_range<-NULL
listingsGLM2$current_delinquencies<-NULL
listingsGLM2$public_records_last10_years<-NULL
listingsGLM2$delinquencies_over60_days<-NULL
listingsGLM2$delinquencies_over90_days<-NULL
listingsGLM2$amount_delinquent<-NULL


logit_group_3<-glm(loan_status_description ~ .,data=listingsGLM2, family = "binomial" )

summary(logit_group_3)

```  

The resulting model that we are left with will be our final model, because it first analyzed all necessary variables in the dataset, and because it has roughly the same AIC, with a lower number of independent variables, thus it likely better explains the data.

```{r}

set.seed(123)
test_set<-sample(1:nrow(listingsGLM2), 8385)
listingsGLM3_train<-listingsGLM2[-test_set,]
listingsGLM3_test<-listingsGLM2[test_set,]

test_predictions_FINAL <- predict(logit_group_3, newdata = listingsGLM3_test, type = "response")


test_predictions_FINAL <- ifelse(test_predictions_FINAL > 0.20, 1, 0)

CrossTable(x = test_predictions_FINAL, y = listingsGLM3_test$loan_status_description, 
           prop.chisq=FALSE)

confusionMatrix(as.factor(test_predictions_FINAL), as.factor(listingsGLM3_test$loan_status_description), positive = "1")

```
  

# Logistic Model Interepretation 

First, interepreting our model. <br/>

The significant variables are: <br/>

- Amount Funded 
- Borrower Rate
- Prosper Rating
- Scorex
- Stated Monthly Income
- Monthly Debt
- Credit Lines in the past 7 years
- Revolving Available Percentage
- Satisfactory Accounts
- Home ownership 

Amount funded and borrower rate are highly significant, and positively correlated to defaults. As they increase, the chance of default increases significantly as well. This makes sense, since a larger loan is likely more difficult to pay back, and a higher interest rate makes that loan harder to pay back as well. 

Prosper rating and scorex are both significant, but only for a few different levels. For prosper rating, only the highest level (AA) is significant. This means that a really high rated borrower would be less likely to default, but after a AA rating, the results are insignificant. The same is true for scorex. Only borrowers with a 748+ scorex rating have a statistically significant lower chance of default than other rating borrowers. This means that borrowers with a 600 rating and borrowers with a 700 rating are not statistically significant from each other. 

Stated monthly income is statistically significant, but the impact is relatively small. However, as stated monthly income increases, the chance of default decreases. The small impact does make sense however, since the value is dollars, and the range of values is many thousands of dollars.

Monthly debt is also very significant, and positively correlated to loan default probability. This means that the more monthly debt that a borrower has, the more likely they are to default on their loans. This makes sense, as more debt likely puts a borrower in a tougher hole to climb out of. 

Credit lines in the past 7 years is also significant, and negatively correlated to default probability. So, this likely means that people that have been offered and given more credit over the past 7 years, are also more likely to pay it back on time. 

Satisfactory accounts is also significant. However, it is positively correlated. This means the more satisfactory accounts that a borrower has, the more likely they are to default. This could also be correlated to monthly debt, as, meaning that a borrower with many satisfactory accounts is also taking on a large amount of debt. 

Finally, home ownership is significant, and POSITIVELY correlated. This means that someone who owns a home is actually more likely to default on a loan than someone who does not. 

# Arbitrage Opportunities 

Here we will look at the differences between the significant factors for the linear and regression models. This analysis will help us understand if the market is missing significant factors correlated to default, or if it is correctly assessing the risks associated with default and correctly pricing loans.

The following variables were significant in our logistic model, but not our regression model:

1. **Monthly Debt:** 
The monthly debt of an individual showcases their financial obligations over a short period of time, which allows a lender to evaluate their debt load compared to their income. Factoring a borrower's monthly debt into the interest rate would better price loans according to their risk profile.

2. **Home ownership:**
Home owners were more likely to default than non-home owners, according to our logistic model. The financial obligation of owning a home and paying a mortgage could potentially impact the ability of an individual to repay a separate loan. This should be taken into account when pricing loans. 

3. **Revolving Available Percentage:** 
The percentage of a loan or credit line a borrower utilizes is an indication of their financial condition and the amount they would have to pay back. If an individual has a revolver fully utilized, they would have to pay back that loan before the new loan. 

## Linear Regression Variables not in Logistic Regression

4. **Income Verifiable:**
Verifying the income of an individual is significant and is a way for lenders to ensure the income of a borrower is what they claim it to be. This was not significant in determining whether a borrower will default. We suspect this is because the vast majority of borrowers had their income verified, and if this was removed it could cause many bad loans. We believe this is an important variable and that it is correctly being used to price loans, despite not being significant in predicting defaults. 

5. **Bank Card Utilization, current credit lines:**
While the utilization rate of an individuals bank card can provide a glimpse into their current borrowing/credit needs, it was not significant in predicting a default. This may be due to the fact that many individuals use a credit card to build credit, earn rewards and track their spending. Also borrowers are using a credit card the way it was meant to be used, and their use may not reflect their actual financial health. This same logic applied to current credit lines.

6. **Installment Balance:**
The installment balance is the regular payment that a borrower makes to the lender, and was not significant in predicting default. While a higher amount provides more risk for the lender due to capital exposure, the borrower may be financially able to manage payments and prior due diligence should prevent unworthy borrowers from receiving large loans. For this reason this variable should not be used to price loans moving forward. 

7. **Loan Status Description:**
The record of ongoing or completed loans was used to price loans, and was negatively correlated with the interest rate. We believe the instance of multiple loans should not be used to price loans since prior due diligence should determine if the borrower is financially capable of taking on an additional loan. 

8. **Employment Status Description:**
The employment status was significant for several levels, however should not be used to price a loan. While the employment status provides a glimpse into the job one holds, it does not predict income, or account for the fact individuals with lower incomes may take out lower loans and therefore not be more likely to default. We recommend removing this criteria when pricing loans. 

9. **Current Delinquencies, Amount Delinquent:**
We were surprised to see this criteria is not significant in predicting default, and believe it should still be used to price loans. Current delinquencies may not have predicted default due to the individual's priority of paying back a particular debt. This is also somewhat uncommon and that may contribute to its muted significance.  We suggest continuing to use this variable when pricing loans.

10. **Public Records (last 12 months), Inquiries (last 6 months), Total Inquiries:**
The inquiries into your finances over a period should not be used to price loans, and may be a product of loan shopping, applying to jobs, etc. While this criteria measures your financial activity to an extent, it does not provide enough information alone to justify increasing the interest rate on a loan. 

## Arbitrage Conclusion

It is clear that this lender is using a set of pre-defined criteria when pricing loans, some of which is not relevant to the borrower's probability of default. The lender should use this information to optimize their pricing strategy to ensure the correct risk-reward profile when issuing loans.

The above analysis of variables shows that we believe several variables should be removed from pricing loans, while some should still be used. The removal of variables that paint an incomplete financial picture of the borrower or unfairly bucket them into groups should be removed, as they are not accurate in predicting default. At the same time many variables should be added to the pricing of loans, including monthly debt, home ownership, and revolving available percentage. 